---
title: "How to produce other key-value uploads"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to produce other key-value uploads}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## When to use this How-To 

Follow the instructions if you want to ...

* Do your own data recoding and aggregation for IPEDS submission 
* Get help converting your data into a key-value pair format
* Get help combining parts with different numbers of columns 


If you want the package to help you with aggregation and (some) recoding, look at the "setup" vignettes for supported uploads.  


## General Process

* **Download/Open the IPEDS Import Spec** for a particular survey and institution-type on the [IPEDS Survey Materials instructions page](https://surveys.nces.ed.gov/ipeds/public/survey-materials/results).
* **Use the "Key Value Pair File" instructions** to prep your data. This will be the second part of the document; the first part is for fixed-width files.
* **Create one dataframe for each survey part** using the exact IPEDS FieldIDs as column names and using only the values described in the table of specs and the additional information in the file's appendix (scroll down).  Follow the same order of columns and parts.
* **Use the package function `produce_other_report`** with your files.

*If you need assistance understanding what goes into the upload file, contact IPEDS for advice.*

## Data prep example: Admissions

This provides a start-to-finish example of preparing an admissions submission based on sample data. Always check your results after you upload your txt file to the IPEDS submission portal,

### Start with institutional data for applicants 
```{r setup}
#load packages
library(dplyr)
library(magrittr)
library(IPEDSuploadables)
```


```{r create_dummy_data, message=FALSE}
#create data
adm_dat <- data.frame(StudentId = seq(1:24),
                      FtPt = c(rep('FT', 23), 'PT'),
                      Sex = rep(c("M", "F"), 12),
                      Admit = c(rep(1, 16), rep(0, 8)),
                      Enroll = c(rep(1, 12), rep(0, 12)),
                      SAT = c(rep(1, 8), rep(0, 16)),
                      SAT_V = c(500, 560, 600, 660, 700, 760, 800, 800, rep(NA, 16)),
                      SAT_M = c(400, 460, 500, 560, 600, 660, 700, 700, rep(NA, 16)),
                      ACT = c(rep(0, 8), rep(1, 16)),
                      ACT_CMP = c(rep(NA, 8), 32, 32, 31, 31, 30, 30, 29, 29, 28, 28, 27, 27, 26, 26, 25, 25)
                      )
```

```{r printdat, echo=FALSE}
knitr::kable(adm_dat, 
            format = 'html')
```

### Prepare Part A - General Questions (not from sample data)
```{r produceA}
#### PART A: General Admissions Criteria
partA <- data.frame(UNITID = 999999,
                    SURVSECT = 'ADM',
                    PART = 'A',
                    ADMCON1 = 2, #GPA
                    ADMCON2 = 1, #Rank
                    ADMCON3 = 1, #Record
                    ADMCON4 = 2, #HS grad
                    ADMCON5 = 1, #Recs
                    ADMCON6 = 3, #Portfolio
                    ADMCON7 = 5, #SAT/ACT  #1 or 5 = have to do part C
                    ADMCON8 = 2, #TOEFL
                    ADMCON9 = 3) #other test
```

```{r printA, echo=FALSE}
knitr::kable(partA, 
            format = 'html')
```
### Prepare Part B - Student Counts (from sample data)
```{r produceB}
##### PART B: Admission Counts; FirstTime UG only
partB <- data.frame(UNITID = 999999,
                    SURVSECT = 'ADM',
                    PART = 'B',
                    APPLCNM = nrow(adm_dat[adm_dat$Sex == 'M', ]),
                    APPLCNW = nrow(adm_dat[adm_dat$Sex == 'F', ]),
                    APPLCNT = nrow(adm_dat),
                    ADMSSNM = nrow(adm_dat[adm_dat$Sex == 'M' & adm_dat$Admit == 1,]),
                    ADMSSNW = nrow(adm_dat[adm_dat$Sex == 'F' & adm_dat$Admit == 1,]),
                    ADMSSNT = nrow(adm_dat[adm_dat$Admit == 1,]),
                    ENRLFTM = nrow(adm_dat[adm_dat$Sex == 'M' & adm_dat$Enroll == 1 & adm_dat$FtPt == 'FT', ]),
                    ENRLFTW = nrow(adm_dat[adm_dat$Sex == 'F' & adm_dat$Enroll == 1 & adm_dat$FtPt == 'FT', ]),
                    ENRLFTT = nrow(adm_dat[adm_dat$Enroll == 1 & adm_dat$FtPt == 'FT', ]),
                    ENRLPTM = nrow(adm_dat[adm_dat$Sex == 'M' & adm_dat$Enroll == 1 & adm_dat$FtPt == 'PT', ]),
                    ENRLPTW = nrow(adm_dat[adm_dat$Sex == 'F' & adm_dat$Enroll == 1 & adm_dat$FtPt == 'PT', ]),
                    ENRLPTT = nrow(adm_dat[adm_dat$Enroll == 1 & adm_dat$FtPt == 'PT', ])
                    )
```

```{r printB, echo=FALSE}
knitr::kable(partB, 
            format = 'html')
```

### Part C: Test Scores (from sample data)
```{r produceC}
#### PART C: Test Scores

adm_enr <- adm_dat %>%
  filter(Enroll == 1)

#we do not supply ACT individual test percentiles, I guess.
partC <- data.frame(UNITID = 999999,
                    SURVSECT = 'ADM',
                    PART = 'C',
                    SATINUM = nrow(adm_enr[adm_enr$SAT == 1, ]),
                    SATIPCT = round(nrow(adm_enr[adm_enr$SAT == 1, ])*100/nrow(adm_enr), 0),
                    ACTNUM = nrow(adm_enr[adm_enr$ACT == 1,]),
                    ACTPCT = round(nrow(adm_enr[adm_enr$ACT == 1,])*100/nrow(adm_enr), 0),
                    SATVR25 = quantile(adm_enr$SAT_V[!is.na(adm_enr$SAT_V)], .25),
                    SATVR75 = quantile(adm_enr$SAT_V[!is.na(adm_enr$SAT_V)], .75),
                    SATMT25 = quantile(adm_enr$SAT_M[!is.na(adm_enr$SAT_M)], .25),
                    SATMT75 = quantile(adm_enr$SAT_M[!is.na(adm_enr$SAT_M)], .75),
                    ACTCM25 = quantile(adm_enr$ACT_CMP[!is.na(adm_enr$ACT_CMP)], .25),
                    ACTCM75 = quantile(adm_enr$ACT_CMP[!is.na(adm_enr$ACT_CMP)], .75),
                    ACTEN25 = -2,
                    ACTEN75 = -2,
                    ACTMT25 = -2,
                    ACTMT75 = -2)
```

```{r printC, echo=FALSE}
knitr::kable(partC, 
            format = 'html', 
            row.names = FALSE)
```



### Use this package to convert those dataframes into a single uploadable txt file

*The file format is shown below, but the package will actually save this as a txt file at the location of your choice.*

```{r producefinal, echo=TRUE, eval=FALSE, message=FALSE}
produce_other_report(partA, partB, partC, survey = "Admissions")
```

```{r showfinal, echo=FALSE, eval=TRUE, message=FALSE}
rbind(IPEDSuploadables:::apply_upload_format(partA),
      IPEDSuploadables:::apply_upload_format(partB),
      IPEDSuploadables:::apply_upload_format(partC)) %>%
  knitr::kable(format = 'html',
               row.names = FALSE,
               col.names = '')
```
### Upload your final txt file to the IPEDS website
This step is no different than any other upload.
