---
title: "Set up your data for IPEDS HR processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Set up your data for IPEDS HR processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE, echo=FALSE}
library(IPEDSuploadables)
```

## To use the `produce_hr_report` function, your data must be in the format described below

>
* One row per person
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)


```{r specstable, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
load("../data/specs_HR.rda")
library(magrittr)
specs_HR$ACCEPTABLE_VALUES <- stringr::str_replace_all(specs_HR$ACCEPTABLE_VALUES, pattern = "\\n", replacement = "<br>")
knitr::kable(specs_HR, 
            format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition')) %>%
    kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```

Notes  
  
* In OccCategory3,  `21` is used for an intermediate total and is not a value in the upload file  
* OccCategory3 values will be recoded in `prep_hr_data_frame` into all necessary rollups, including instructional staff (values 1-4)

