---
title: "Set up your data for IPEDS 12 Month Enrollment processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Set up your data for IPEDS 12 Month Enrollment processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE, echo=FALSE}
library(IPEDSuploadables)
```

## To use the `produce_e1d_report` function, you must provide two (2) dataframes: one with student enrollment data and one with instructional activity data.

### The student enrollment data must be in the format described below

>
* One row per person
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)

Note: Column names can use any capitalization method you like

```{r create_data_file_in_package, eval=FALSE, echo=FALSE}
#this chunk is only used during interactive session by the package maintainers, if the requirements have been updated

tryCatch(specs_E1D <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1bhjgAjmipg_pqwUum2Wcn6sOKInoTUne_PeFDKf3ncs/edit", sheet = 1),
        error = function(x) message("Google Sheet unavailable"),
        warning = function(y) message("There was a warning"))

usethis::use_data(specs_E1D, overwrite = TRUE)
```

```{r specstable_df, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
data(specs_E1D, package = "IPEDSuploadables")
library(magrittr)
specs_E1D$ACCEPTABLE_VALUES <- stringr::str_replace_all(specs_E1D$ACCEPTABLE_VALUES, pattern = "\\n", replacement = "<br>")
specs_E1D_startingdf <- specs_E1D[specs_E1D$DATA_FRAME == 'startingdf', 2:4]
knitr::kable(specs_E1D_startingdf,
             format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition')) %>%
  kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```



### The instructional activity data must be in the format described below

>
* One row total, with a column for each reportable fact
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)


```{r specstable_e1d, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
specs_E1D_hrs <- specs_E1D[specs_E1D$DATA_FRAME == 'hrs', 2:4]
knitr::kable(specs_E1D_hrs,
             format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition'),
             row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```


