---
title: "Set up your data for IPEDS Completions processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Set up your data for IPEDS Completions processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE, echo=FALSE}
library(IPEDSuploadables)
```

## To use the `produce_com_report` function, your student completions data must be in the format described below

>
* One row per person per degree level per cip per major
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)

Note: Column names can use any capitalization method you like

```{r create_data_file_in_package, eval=FALSE, echo=FALSE}
#this chunk is only used during interactive session by the package maintainers, if the requirements have been updated

tryCatch(specs_COM <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/132EKp46TpEHPlVrG9FZxINxmVVz8UO0Ov989sP4lNaP/edit", sheet = 1), 
        error = function(x) message("Google Sheet unavailable"),
        warning = function(y) message("There was a warning"))

usethis::use_data(specs_COM, overwrite = TRUE)

#alternatively, this code combination works for save/load and doesn't end up as package data
#write.csv(x = specs_COM, file = "../vignettes/specs_COM.csv", row.names = FALSE)
#specs_COM <- data.frame(read.csv("specs_COM.csv", stringsAsFactors = FALSE))
# but the file size for the csv is slightly larger than the rda
```

```{r specstable_df, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
data(specs_COM, package = "IPEDSuploadables")
library(magrittr)
specs_COM$ACCEPTABLE_VALUES <- stringr::str_replace_all(specs_COM$ACCEPTABLE_VALUES, pattern = "\\n", replacement = "<br>")
specs_COM_startingdf <- specs_COM[specs_COM$DATA_FRAME == 'startingdf', 2:4]
knitr::kable(specs_COM_startingdf,
             format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition')) %>%
    kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```



### If there are any cip code/degree level/major number combinations that are possible for students to obtain but are NOT represented in this year's student completions data, use the format below to incorporate them into the `make_com` functions.

>
* One row per degree level per cip per major
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)


```{r specstable_com, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
specs_COM_extracips <- specs_COM[specs_COM$DATA_FRAME == 'extracips', 2:4]
knitr::kable(specs_COM_extracips,
             format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition'),
             row.names = FALSE) %>%
    kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```


