---
title: "Set up your data for IPEDS Completions processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Set up your data for IPEDS Completions processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE, echo=FALSE}
library(IPEDSuploadables)
```

## To use the `make_com` functions, your student completions data must be in the format described below

>
* One row per person per degree level per cip per major
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)


```{r create_data_file_in_package, eval=FALSE, echo=FALSE}
#export spreadsheet with specs from google drive
#read in with readr
#then run: use_data(specs_COM, specs_COM) to save as .rda
```

```{r specstable_df, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
load("../data/specs_COM.rda")
library(magrittr)
specs_COM$ACCEPTABLE_VALUES <- stringr::str_replace_all(specs_COM$ACCEPTABLE_VALUES, pattern = "\\n", replacement = "<br>")
specs_COM_startingdf <- specs_COM[specs_COM$DATA_FRAME == 'startingdf', 2:4]
knitr::kable(specs_COM_startingdf,
             format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition')) %>%
    kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```



### If there are any cip code/degree level/major number combinations that are possible for students to obtain but are NOT represented in this year's student completions data, use the format below to incorporate them into the `make_com` functions.

>
* One row per degree level per cip per major
* One IPEDS Unitid per file
* Columns with values and types as described below (additional columns are allowed)


```{r specstable_com, echo=FALSE, warning=FALSE, message=FALSE}
#figure out why striped isn't working when format = 'html'
specs_COM_extracips <- specs_COM[specs_COM$DATA_FRAME == 'extracips', 2:4]
knitr::kable(specs_COM_startingdf,
             format = 'html',
             escape = FALSE,
             col.names = c('Column Name', 'Column Type', 'Acceptable Value - Definition')) %>%
    kableExtra::kable_styling(bootstrap_options = "striped") %>%
  kableExtra::column_spec(2, italic = TRUE, width = '8em')

```


