---
title: "How to produce an IPEDS Completions upload file"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to produce an IPEDS Completions upload file}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Step 0: Install/Load the package
```{r setup, eval=FALSE}
#install option 1
devtools::install_github("AlisonLanski/IPEDSuploadables")

#install option 2
githubinstall::githubinstall("IPEDSuploadables")

#load the packages
library(IPEDSuploadables)
```

## Step 1: Load your institution data

Read in your data 
For setup requirements, consult [this prep vignette](setup_for_completions.html) for column names, column types, and accepted values.

```{r prep_data, eval = FALSE}
dat_student <- readr::read_csv::("path/StudentData.csv")
dat_cips <- readr::read_csv::("path/CipData.csv")
```


To test the functions without using your own institutional data, you can create dummy data by running  
*Note: 1 row will fail to parse in the dummy data: this is an intentionally-incorrect birthdate and does NOT cause anything later to fail*

```{r dummy_data, eval = FALSE}
dat_student <- create_dummy_data_com()
dat_cips <- create_dummy_data_com("extracips")
```


## Step 2: Produce your IPEDS report

Run the wrapper function to produce all portions of your report into a txt for upload.  

* Only the first argument (df) is required and is for your student data.  
* To run only a single part of the report, change "ALL" to "A", "B", "C", or "D"
* The first time you run this function, you'll be asked to browse to the file location where the report will be saved. This location can be changed later.

```{r produce_report, eval=FALSE}
#full export
produce_com_report(df = dat_student, extracips = dat_cips, part = "ALL")

#if you have no extracips and want to look at part B
produce_com_report(dat_student, part = "B")
```


This file will produce a status message for each part that is asked for.  When running "all" completions parts, you'll see the message 4 times even though there is only 1 output file.  This is a known bug that doesn't affect functionality.

#### Note: If you prefer to avoid the wrapper function, you can run its individual parts.  
First, prep the data by running it through a script which checks for cip-code formatting and regularizes some datatypes  
*Prep script is minimal right now; we have plans for more robust checks*
Then, run the individual component scripts
Use the help for these components to see how to produce the txt file, an easier-to-view file, or both.

```{r prep_script, eval = FALSE}
dat_student <- prep_com_data_frame(dat_student)
dat_cips <- prep_com_data_frame(dat_cips)
```

```{r part_script, eval = FALSE}
make_com_part_A(dat_student, extracips)
make_com_part_B(dat_student, extracips)
make_com_part_C(dat_student, extracips)
make_com_part_D(dat_student_extracips)
```

We recommend only using the individual component scripts for troubleshooting.  To submit the individual results in Step 3, you'll need to first combine them into a single text file.


## Step 3: Upload your IPEDS report

A. Navigate to the IPEDS submission portal and log in.  
B. Select "Completions", select "key-value pair", then browse to your .txt file and upload.

* If you receive errors about missing CIPS reported last year, add those values to your extracips file.  
* If you receive errors about duplicate keys, check that your extracips file does not have cips which are also part of your student file.  
  
C. After your file uploads, continue your standard process to check edits, enter metadata, review results, lock, and submit the survey.

